---
title: "Animating the DeZIM Institut Logo"
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [DeZIM](https://www.dezim-institut.de) (Deutsches Zentrum für
Integration- und Migrationsforschung) logo consists of four overlapping
circles with Venn-style colour blending, five white letterforms, a vertical
divider, and a typographic block to the right. The animation reveals each
element in a choreographed sequence: the four circles scale in left to right,
the shared lens regions fade in, the letterforms follow with a stagger, the
divider grows from its midpoint, and the text settles in from a slight
rightward offset.

## 1. SVG annotation

The original SVG carries no animation identifiers. The code below uses `sub()`
to inject `data-animejs-id` attributes on the five individually targeted
elements and to prepend initial-state CSS rules to the existing
`<defs><style>` block. Elements animated as groups are reached via class
selectors.

```{r annotate}
library(animejs)

svg_raw <- paste0(
  readLines(
    "https://www.dezim-institut.de/_assets/03ba20435fb0f29076b2967fa9a28b23/Images/logo-dezim.svg",
    warn = FALSE
  ),
  collapse = "\n"
)

svg_annotated <- svg_raw |>
  sub(
    pattern = "<defs><style>",
    replacement = paste0(
      "<defs><style>",
      ".anim-circle{opacity:0;transform:scale(0)}",
      ".anim-divider{opacity:0;transform:scaleY(0)}",
      ".e{opacity:0}",
      ".f,.j,.l{opacity:0}",
      ".i{opacity:0;transform:translateX(10px)}"
    ),
    fixed = TRUE
  ) |>
  sub(
    pattern = '<path class="d"',
    replacement = '<path data-animejs-id="circle-teal" class="d anim-circle" style="transform-origin:49.5px 49.5px"',
    fixed = TRUE
  ) |>
  sub(
    pattern = '<path class="h"',
    replacement = '<path data-animejs-id="circle-amber" class="h anim-circle" style="transform-origin:89.1px 49.5px"',
    fixed = TRUE
  ) |>
  sub(
    pattern = '<path class="k"',
    replacement = '<path data-animejs-id="circle-red" class="k anim-circle" style="transform-origin:148.5px 49.5px"',
    fixed = TRUE
  ) |>
  sub(
    pattern = '<path class="g" d="M234',
    replacement = '<path data-animejs-id="circle-orange" class="g anim-circle" style="transform-origin:218px 49.5px" d="M234',
    fixed = TRUE
  ) |>
  sub(
    pattern = '<path class="g" d="M253',
    replacement = '<path data-animejs-id="divider" class="g anim-divider" style="transform-origin:253.15px 49.5px" d="M253',
    fixed = TRUE
  )
```

## 2. Timeline

```{r timeline}
tl <- anime_timeline() |>
  # Four circles scale in from their geometric centres, left to right.
  # anime_easing_back() produces a brief overshoot that reinforces the
  # circular geometry of the icon.
  anime_add(
    selector = anime_target_id("circle-teal"),
    props = list(opacity = anime_from_to(0, 1), scale = anime_from_to(0, 1)),
    duration = 700,
    ease = anime_easing_back("out"),
    offset = 0
  ) |>
  anime_add(
    selector = anime_target_id("circle-amber"),
    props = list(opacity = anime_from_to(0, 1), scale = anime_from_to(0, 1)),
    duration = 700,
    ease = anime_easing_back("out"),
    offset = 200
  ) |>
  anime_add(
    selector = anime_target_id("circle-red"),
    props = list(opacity = anime_from_to(0, 1), scale = anime_from_to(0, 1)),
    duration = 700,
    ease = anime_easing_back("out"),
    offset = 400
  ) |>
  anime_add(
    selector = anime_target_id("circle-orange"),
    props = list(opacity = anime_from_to(0, 1), scale = anime_from_to(0, 1)),
    duration = 700,
    ease = anime_easing_back("out"),
    offset = 600
  ) |>
  # Venn lens regions fade in once all four circles are present.
  anime_add(
    selector = anime_target_css(".f, .j, .l"),
    props = list(opacity = anime_from_to(0, 1)),
    duration = 400,
    ease = anime_easing("Quad", "out"),
    offset = 900
  ) |>
  # White letterforms (D, E, Z, I, M) fade in left to right with a stagger.
  # The five .e paths appear in left-to-right DOM order, so the default
  # stagger direction ("first") matches the reading direction of the letters.
  anime_add(
    selector = anime_target_css(".e"),
    props = list(opacity = anime_from_to(0, 1)),
    duration = 400,
    ease = anime_easing("Quad", "out"),
    stagger = anime_stagger(80),
    offset = 1050
  ) |>
  # Divider line grows from its vertical midpoint.
  anime_add(
    selector = anime_target_id("divider"),
    props = list(opacity = anime_from_to(0, 1), scaleY = anime_from_to(0, 1)),
    duration = 500,
    ease = anime_easing("Expo", "out"),
    offset = 1300
  ) |>
  # Text block fades in and settles leftward from a 10 px rightward offset.
  anime_add(
    selector = anime_target_css(".i"),
    props = list(
      opacity = anime_from_to(0, 1),
      translateX = anime_from_to(10, 0)
    ),
    duration = 700,
    ease = anime_easing("Quad", "out"),
    offset = 1600
  )
```

## 3. Render

```{r render}
tl |>
  anime_playback(autoplay = TRUE, loop = TRUE, controls = TRUE) |>
  anime_render(svg_annotated, width = 520, height = 120)
```
