---
title: "Animating a Rowing Stroke with animejs"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The rowing stroke

A complete stroke has two phases.

**Drive** (fast, ~800 ms): starting from the *catch* -- knees bent, body
forward, arms extended toward the flywheel -- the rower pushes with the legs,
swings the body back, and pulls the handle to the chest. The seat slides away
from the flywheel.

**Recovery** (slower, ~900 ms): arms extend first, then the body rocks
forward, and finally the legs bend to return the seat to the catch position.

The animation loops these two phases continuously.

## 1. The machine frame

The machine is made of static SVG elements. The flywheel sits on the left; a
monorail extends to the right; a foot stretcher is fixed partway along the
rail.

```{r machine-svg}
machine_frame <- r"(
  <!-- Rail -->
  <line x1="58" y1="175" x2="400" y2="175"
        stroke="#64748b" stroke-width="4" stroke-linecap="round"/>

  <!-- Rail support legs -->
  <line x1="120" y1="175" x2="120" y2="192"
        stroke="#64748b" stroke-width="3"/>
  <line x1="340" y1="175" x2="340" y2="192"
        stroke="#64748b" stroke-width="3"/>

  <!-- Flywheel housing -->
  <circle cx="88" cy="150" r="30"
          fill="none" stroke="#64748b" stroke-width="3.5"/>

  <!-- Flywheel spokes -->
  <line x1="88" y1="120" x2="88" y2="180"
        stroke="#94a3b8" stroke-width="1.5"/>
  <line x1="58" y1="150" x2="118" y2="150"
        stroke="#94a3b8" stroke-width="1.5"/>
  <line x1="67" y1="129" x2="109" y2="171"
        stroke="#94a3b8" stroke-width="1.5"/>
  <line x1="109" y1="129" x2="67"  y2="171"
        stroke="#94a3b8" stroke-width="1.5"/>

  <!-- Foot stretcher -->
  <line x1="192" y1="157" x2="196" y2="175"
        stroke="#64748b" stroke-width="3.5" stroke-linecap="round"/>
  <line x1="184" y1="157" x2="200" y2="157"
        stroke="#64748b" stroke-width="3.5" stroke-linecap="round"/>
)"
```

## 2. The rower

All rower elements are drawn at the **catch** position. This is the baseline
from which animejs computes transitions; the SVG coordinates encode catch
geometry directly, so no initial CSS transform is required on any element.

The rower is broken into six elements. Each carries a `data-animejs-id`
attribute so the R timeline can target it by a selector produced by
`anime_target_id()`.

### 2.1 Seat

The seat is a rounded rectangle that slides 95 px to the right during the
drive (the leg push moves the seat away from the flywheel).

```{r seat-svg}
seat_svg <- r"(
  <rect data-animejs-id="seat"
        x="205" y="165" width="28" height="10" rx="3"
        fill="#475569"/>
)"
```

### 2.2 Leg

A single line from the fixed foot (192, 175) to the hip at catch (219, 165)
represents the leg. Rather than a fully articulated shin and thigh, this
schematic line communicates the leg extension clearly. Because animejs can
animate SVG presentation attributes directly, the hip end of the line is
animated by updating `x2` from 219 (catch) to 314 (finish) as the seat
slides.

```{r leg-svg}
leg_svg <- r"(
  <line data-animejs-id="leg"
        x1="192" y1="175" x2="219" y2="165"
        stroke="#1e293b" stroke-width="3.5" stroke-linecap="round"/>
)"
```

### 2.3 Torso

The torso is a 60 px line from the hip (219, 165) to the shoulder (193, 111).
As drawn, it leans 26° forward from vertical -- the catch body angle. The
`transform-origin` is pinned to the hip so that all CSS rotation pivots there.

During the drive the hip translates 95 px right (`translateX`) and the body
swings 41° clockwise (`rotate`), which brings it from −26° to +15° relative
to vertical (a back lean at the finish). The head shares the same pivot and
the same transform values.

```{r torso-svg}
torso_svg <- r"(
  <line data-animejs-id="torso"
        x1="219" y1="165" x2="193" y2="111"
        stroke="#1e293b" stroke-width="4" stroke-linecap="round"
        style="transform-origin: 219px 165px"/>
)"
```

### 2.4 Head

The head is a circle centred 12 px above the shoulder at catch (190, 98).
Like the torso, it pivots around the hip (219, 165) and receives identical
`translateX` and `rotate` values, so it tracks the top of the torso exactly.

```{r head-svg}
head_svg <- r"(
  <circle data-animejs-id="head"
          cx="190" cy="98" r="12"
          fill="#1e293b"
          style="transform-origin: 219px 165px"/>
)"
```

### 2.5 Arms

The arm is a 50 px line from the shoulder (193, 111) to the hands (150, 135),
which points toward the flywheel handle at the catch. At the finish the
shoulder has moved to (330, 107) and the arms are pulled straight down to hip
level.

Three transforms compose the drive motion:

- `rotate(-61)` -- swings the arm counterclockwise from the flywheel-reaching
  angle (~151° from horizontal) to pointing straight down (90°), pivoting at
  the catch shoulder position.
- `translateX(137)` -- carries the shoulder from x = 193 to x = 330 as the
  torso moves with the seat.
- `translateY(-4)` -- accounts for the small vertical rise of the shoulder as
  the torso rotates back.

CSS applies these right to left: rotate first, then translateY, then
translateX. The net effect is that the shoulder lands precisely on the torso
top at the finish.

```{r arms-svg}
arms_svg <- r"(
  <line data-animejs-id="arms"
        x1="193" y1="111" x2="150" y2="135"
        stroke="#1e293b" stroke-width="3.5" stroke-linecap="round"
        style="transform-origin: 193px 111px"/>
)"
```

### 2.6 Handle

The handle is a dashed line from the flywheel edge (118, 150) to the hands
(150, 135) at catch. Like the leg, it is animated via SVG attribute animation:
`x2` and `y2` track the hands from catch to finish and back.

```{r handle-svg}
handle_svg <- r"(
  <line data-animejs-id="handle"
        x1="118" y1="150" x2="150" y2="135"
        stroke="#94a3b8" stroke-width="2.5"
        stroke-dasharray="5 3" stroke-linecap="round"/>
)"
```

### 2.7 Assembling the SVG
```{r full-svg}
library(animejs)

svg_src <- paste0(
  '<svg viewBox="0 0 440 210" xmlns="http://www.w3.org/2000/svg">',
  machine_frame,
  seat_svg,
  leg_svg,
  torso_svg,
  head_svg,
  arms_svg,
  handle_svg,
  '</svg>'
)
```

## 3. Building the timeline

The timeline has two segments per loop: the drive and the recovery. Both
segments target all six animated elements. Elements that use CSS transforms
(`seat`, `torso`, `head`, `arms`) are targeted by `anime_target_id()`;
elements animated via SVG attributes (`leg`, `handle`) receive separate
`anime_add()` calls with `x2`/`y2` props.

### 3.1 Drive phase (catch → finish)

The elements drawn at catch require no explicit `from` value -- animejs reads
their current state from the DOM. Only the target (finish) values are
specified. The drive uses `anime_easing("Quart", "out")` for a forceful
initial push that decelerates toward the finish.

```{r drive}
tl <- anime_timeline(
  duration = 800,
  ease = anime_easing("Quart", "out"),
  loop = TRUE
) |>
  # Seat slides right
  anime_add(
    selector = anime_target_id("seat"),
    props = list(translateX = 95)
  ) |>
  # Hip end of leg moves right with the seat
  anime_add(
    selector = anime_target_id("leg"),
    props = list(x2 = 314),
    offset = 0 # starts at t = 0, same as seat
  ) |>
  # Torso: hip moves right, body swings to back lean
  anime_add(
    selector = anime_target_id("torso"),
    props = list(translateX = 95, rotate = 41),
    offset = 0
  ) |>
  # Head: identical transforms to torso (shares hip pivot)
  anime_add(
    selector = anime_target_id("head"),
    props = list(translateX = 95, rotate = 41),
    offset = 0
  ) |>
  # Arms: rotate from flywheel-reach to pulled-in, shoulder tracks torso top
  anime_add(
    selector = anime_target_id("arms"),
    props = list(
      translateX = 137,
      translateY = -4,
      rotate = -61
    ),
    offset = 0
  ) |>
  # Handle tracks the hands
  anime_add(
    selector = anime_target_id("handle"),
    props = list(x2 = 330, y2 = 156),
    offset = 0
  )
```

### 3.2 Recovery phase (finish → catch)

The recovery uses `anime_from_to()` throughout to explicitly set both the
finish value (from) and the catch value (to), since the DOM is now in the
finish state. The recovery easing (`"inOut", "Sine"`) is smooth and
symmetrical, reflecting the controlled, unhurried return.

The recovery starts 100 ms after the drive ends, giving a brief pause at the
finish position.

```{r recovery}
tl <- tl |>
  anime_add(
    selector = anime_target_id("seat"),
    props = list(translateX = anime_from_to(95, 0)),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = "+=100"
  ) |>
  anime_add(
    selector = anime_target_id("leg"),
    props = list(x2 = anime_from_to(314, 219)),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = 900
  ) |>
  anime_add(
    selector = anime_target_id("torso"),
    props = list(
      translateX = anime_from_to(95, 0),
      rotate = anime_from_to(41, 0)
    ),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = 900
  ) |>
  anime_add(
    selector = anime_target_id("head"),
    props = list(
      translateX = anime_from_to(95, 0),
      rotate = anime_from_to(41, 0)
    ),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = 900
  ) |>
  anime_add(
    selector = anime_target_id("arms"),
    props = list(
      translateX = anime_from_to(137, 0),
      translateY = anime_from_to(-4, 0),
      rotate = anime_from_to(-61, 0)
    ),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = 900
  ) |>
  anime_add(
    selector = anime_target_id("handle"),
    props = list(
      x2 = anime_from_to(330, 150),
      y2 = anime_from_to(156, 135)
    ),
    duration = 900,
    ease = anime_easing("Sine", "inOut"),
    offset = 900
  )
```

## 4. Rendering

```{r render}
tl |>
  anime_playback(controls = TRUE) |>
  anime_render(svg = svg_src, width = "360px")
```
